---
title: "analysis"
output: html_document
---

# Analysis of Final Year Project Data

## Data Preperation 

```{r load packages, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Loading packages

library(tidyverse)
library(here)
library(aplpack)
library(interactions)
library(car)
library(lmtest)
library(stats)

```


```{r load in data, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Load in the data

survey <- read.csv(here("questionnaires.csv"))

task <- read.csv(here("navon_task.csv"))

```

```{r selected variables, warning = FALSE,echo = FALSE, message=FALSE, results= 'hide'}

## Selecting required variables

survey <- survey %>% 
                 select(ResponseId, 22:70)

task <- task %>%
            select(Participant.Public.ID, Zone.Type, 
                   Response, Correct, attention,practice_attention,
                   Reaction.Time, conditions, practice_conditions)

```

Loaded the data into Rstudio and selected the variables required for statistical analysis.

## Data Wrangling

### Dataset Modification 

```{r selected variables, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Formatting existing variables

## Select response rows
task <- task %>%
            subset(Zone.Type == "response_keyboard_single")

## Combine test and practice sections due to randomization error
task$instructed_attentional_focus <- paste(task$attention, task$practice_attention) # Attentional focus condition

task$stimuli <- paste(task$conditions, task$practice_conditions) # Stimuli presented


## Clean up dataframe 
task <- task %>%
            select(-Zone.Type) # remove column

task <- task %>%
           select(-attention, -practice_attention,
                  -conditions, - practice_conditions) # remove columns

survey <- survey[-c(1,2),] #remove unnecessary rows


```

Selected specific response rows in the dataframe that is needed for statistical analysis. The randomisation procedure in the task randomised the order of all trials rather than randomising the order of trials per block resulting in the denoted practice trials occurring at different temporal periods. To deal with this issue the test and practice sections labeled in the data was first combined. Any unnecessary rows and columns that is not needed for analysis was removed from the dataframe to improve readability. 


```{r merging datasets, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Rename columns

colnames(task)[1] <- "id"
colnames(task)[2] <- "response"
colnames(task)[3] <- "correct"
colnames(task)[4] <- "response_time"

colnames(survey)[1] <- "id"

## Merge Dataframes

df <- inner_join(task, survey, "id")

```

To help with readability and coding, some variables were renamed. The separate dataframes were merged by the function inner_join() to remove participants that did complete both the task and survey. 

### Create New Identification Variables

#### Trial Order and Blocks
```{r new variables, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Create numeric ID
df <- df %>% group_by(id) %>% 
              mutate(id_num=cur_group_id()) 

## Create trial number per participant
df <- df %>% group_by(id) %>%
              mutate(trial_num= row_number()) 

## Code practice trials
df <- df %>% mutate(trial_type = cut(trial_num, breaks=c(-Inf, 6.1, Inf), 
                                     labels=c("practice","test"))) 

## Code blocks of trails
df <- df %>% mutate(block = cut(trial_num, breaks = c(0, 6.1, 18.1, 30.1,
                                                      42.1, 54.1, 66.1,78.1),
                    labels = c("practice", "block_1", "block_2", "block_3",
                             "block_4", "block_5", "block_6"))) 

## Delete practice trial rows

df <- df %>%
        subset(trial_type == "test") # Remove unnecessary responses

## Remove trial_type column
df <- df %>%
            select(-trial_type) 

```

A numeric ID was created for easier readability. The temporal order of each trail is labeled per participant. The block number of the trials were labeled and practice trials were removed from the dataframe. 


#### Switching Type per Trials and Previous Errors
```{r switching coding, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Specify function from dplyr package
recode <- dplyr::recode

## Recode attention focus variable to numeric variable
df <- df %>%
  mutate(attention_num = recode(instructed_attentional_focus,
                             "SMALL.jpg " = 1, " SMALL.jpg" = 1,
                             "LARGE.jpg "= 2, " LARGE.jpg" = 2)) 

## Create switching variables for each trial 
## Based upon previous trial focus
df<- df %>% group_by(id, block) %>% 
           mutate(switch_type = 
           if_else(lag(attention_num) == attention_num, "no_switch",
           if_else((lag(attention_num) == 1 & attention_num == 2),"local_global",
           if_else((lag(attention_num) == 2 & attention_num == 1),"global_local", "other")))) # Create categories


## Code trials in which previous trial had error
## Delete response time of rows in which previous trial has error
df <- df %>% group_by(id,block) %>%
         mutate(previous_error=
         if_else(lag(correct) == 0, "yes", "no"))

## Make NA from first trials in block equal no
df$previous_error[is.na(df$previous_error)] <- "no"

```

A switch type variable was created to denote whether the trial represents a switch from a different attentional focus. This variable was created within blocks as it is not known whether participants took a break after completing a block of trials. A previous error variable was created to denote whether the previous trial contained an error response, this was created to capture unsuccessful attentional switching for later analyses.   

#### Accuracy percentage

```{r accuracy, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Calculate accuracy percentage 
df <- df %>% 
  group_by(id) %>% 
  mutate(accuracy = sum(correct)/n()*100)

## See accuracy per participant
acc <- df %>% select(id_num, id, accuracy) %>%
              distinct(id, .keep_all = TRUE) %>% 
              arrange(accuracy)

```

Create accuracy percentage for each participant.

### Recode Survey Data 
#### Autistic Traits
```{r trait to numerical, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Temporarily load in package for recode process
library(plyr)

## Create recode function for traits
recode_trait <- function(x) {
  #recodes autistic traits items to numerical data
  revalue(x, c('Definitely Agree' = 1, 'Agree'= 2, 'Disagree' = 3,
               'Definitely Disagree' = 4)) } 

## Apply function on all trait variables
df[,9:20] <- lapply(df[,9:20], recode_trait)

```

```{r reverse code traits, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Make dataframe without reverse coding for checks later
dfnrc <- df

## Trait items 4, 6, 7 and 12 reverse coding
## Create reverse coding function
reverse_code <- function(x) {
  # Reverse codes for items with 4 scales
  revalue(x, c( "1" = 4, "2" = 3, "3" = 2, "4" = 1)) } 

## Apply function on all bully variables
df[c("TRAIT_4", "TRAIT_6", 
     "TRAIT_7", "TRAIT_12")] <- lapply(df[c("TRAIT_4", "TRAIT_6",
                                            "TRAIT_7", "TRAIT_12")],
                                          reverse_code)

## Change from traits from categorical to numeric variable
df[,9:20] <- lapply(df[,9:20], as.numeric)
dfnrc[,9:20] <- lapply(dfnrc[,9:20], as.numeric)



```


#### Psychotic-Like Experiences
```{r recode ple data, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Create recode function for ple
recode_ple <- function(x) {
  #recodes ple items to numerical data
  revalue(x, c("Nearly Always" = 3, "Often" = 2, 
               "Sometimes" = 1, "Never" =0)) } 


## Apply function on all ple variables
df[,21:35] <- lapply(df[,21:35], recode_ple)

## Change from categorical to numeric variable
df[,21:35] <- lapply(df[,21:35], as.numeric)

## Change age to numeric
df[,8] <- lapply(df[,8],as.numeric)


```

#### Bully Scales
```{r recode bully data, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Create recode function for bully
recode_bully <- function(x) {
  #recodes bully experience items to numerical data
  revalue(x, c("Occurred several times a week" = 4, 
               "Took place once a week" = 3,
               "Occurred two or three times a month" = 2, 
               "Took place once a month" = 1,
               "Did not occur" = 0)) } 


## Apply function on all bully variables
df[,36:54] <- lapply(df[,36:54], recode_bully)

## Change from categorical to numeric variable
df[,36:54] <- lapply(df[,36:54], as.numeric)

## Remove package to avoid masking effects
detach(package:plyr)
```

## Data Cleaning

### Trial Completion

```{r check trials, warning = FALSE, echo= FALSE , message=FALSE,results= 'hide'}

## See max trials per ppt
trial <- aggregate(df$trial_num, by = list(df$id_num), max) 

## See pppt who did not complete task
trial %>% filter(x< 78) # ppt num 80 & 119 only completed 18 and 12 trials
                        

## Remove participants 80 & 199 due to insuffcient data
df <-df[!(df$id_num == "80" | df$id_num =="119"),] 

```

Participants 80 and 119 were deleted from the analysis as after deducting the 6 practice trials they only have data from 12 and 6 trials respectively. All of the other participants completed all 78 trials. 


### Dealing with Missing Data 

```{r missing data, echo=FALSE, fig.show='hide', message=FALSE, warning=FALSE, results='hide'}

## Make dataframe with unique observations for survey data
na <- df %>%  
         group_by(id_num) %>%
         distinct(id_num, .keep_all = TRUE) 

## Check amount of missing data
summary(na)

## Create function to measure percent of missing data
pMiss <- function(x){sum(is.na(x))/length(x)*100}

apply(na,2,pMiss) # Use function on columns
apply(na,1,pMiss) # Use function on rows

## Delete participant 138
df <- df[!(df$id_num == "138"),]

## Tempararily load packages
library(mice)

## Using predictive mean matching to impute missing data
tempdata <- mice(df,m=5,maxit=50,meth='pmm',seed=500)

## Convert into dataframe
df <- complete(tempdata,1)

## Remove package to avoid masking effects
detach(package:mice)

```

Missing data is only present in the traditional bullying victimisation scale, this effect is likely due to the survey forcing a response from participants on other questionnaires. The number of missing data points range from 1 to 4 in each item. No item has over 5% missing data with the highest missing data being 1.4% in item 5. Participant number 138 did not answer any question on the traditional bullying victimisation scale. As such, participant 138 was listwise deleted from the analysis so that bias is not introduced from imputations. The remaining missing values were imputed using a predictive mean matching imputation approach. Only items 1, 3 and 6 had missing values following the deletion of participant 138.

### Checking for Participants Compliance

#### Congruency Errors with No switching

```{r errors, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Create variable to detect congruent errors
df5 <- df %>%  mutate(congruent_error =
         if_else(correct == 0 & stimuli == "congruent_h.jpg ", "1",
         if_else(correct == 0 & stimuli == " congruent_h.jpg", "1",
         if_else(correct == 0 & stimuli == " congruent_s.jpg", "1", 
         if_else(correct == 0 & stimuli == "congruent_s.jpg ", "1", "0")))))


## Show only congruent errors
df5 <- df5 %>% filter(congruent_error == "1" & switch_type == "no_switch") # 232 congruent errors in data set

## Change from categorical to numeric variable
df5[,63] <- lapply(df5[,63], as.numeric) 


## Sum congruent errors with no switch per ppts
df5 <- df5 %>% group_by(id_num) %>%
           mutate(sum_no_switch_con_error = sum(congruent_error))

## Check amount of errors below 60% accuracy 
df6 <- df5 %>% select(id_num, response_time, accuracy, congruent_error, sum_no_switch_con_error, trial_num)  

df6 <- df6 %>% distinct(id_num, .keep_all = TRUE)


## Visualisation of errors and accuracy
individual_plot <- ggplot(data = df6, aes(sum_no_switch_con_error, accuracy))+geom_line()+geom_point()+ scale_x_continuous(breaks= seq(0, 10, 1))

print(individual_plot)
## 4 and above errors appears to be out of character
## 164, 114, 15, 140, 156, 228, 5, 202, 151, 29, 174, 257, 53


## will check if at beginning of trial order
df7 <- df5 %>% filter(id_num == 53)

individual_plot <- ggplot(data = df7, aes(trial_num, sum_no_switch_con_error))+geom_line()+geom_point()+ scale_x_continuous(breaks= seq(0, 80, 10))

print(individual_plot)


```

To gain an idea of participants that were compliant at the task instructions and were concentrating on the task, the amount of errors in congruent trails that did not require a switch in attention. This condition was the easiest so larger errors will indicate poorer compliance, a lack of concentration or a misunderstanding of the task instructions. 113 participants made at least one of these kinds of errors. When plotting accuracy in performance and the total of congruent no switching errors per participants 4 and more errors appear problematic. When these errors occurred in the task was visualised per participant and the following trends appeared. Participant number 164, 114, 15, 140, 156, 5, 202, 257 errors appeared throughout the temporal order of the trail. Whereas, 174 appeared near the beginning and end, 151 occurred near the end, 29, 228 and 53 occurred near the beginning. These participants will be considered for deletion whilst testing further issues. 

#### Response Order

```{r response, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide', eval = FALSE}


## Graphics of response order

## Create list for iteration
id_num  <- c(1:280)

## Make response plot per trail for each participant 
for (i in unique(id_num)){
  
  d <- subset(df, id_num == i)
  
plot <- ggplot(data = d, aes(trial_num, response, group = id_num, colour = block)) + geom_line()+geom_point()+ ggtitle(paste("ID", i))

print(plot)
  
}


```

A plot of each participants response over trials was created to identify any responses that demonstrate task non-compliance such as only one response in general or one response in each block. The analysis revealed that participant 257 only responded with "s" throughout  the task. It is difficult to determine if other responses are genuine from the plots. Participant 257 will be removed from the analysis 

#### Response Switching

```{r response switch, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}


## Remove partcitipant

df <- df[!(df$id_num == "257"),]


## Response switching 

## Create numerical response variable
dfs <- df %>% mutate(num_response = if_else(response == "s", "1", if_else(response == "h", "0", "NA")))

## Calcuate if response switch across trial occured
df2 <- dfs %>% group_by(id) %>%
       mutate(response_switch = ifelse(num_response == lag(num_response), "0", "1"))

## Make numeric variable
df2[,64] <- lapply(df2[,64], as.numeric) 

## Add each participants total number of switching
df3 <- df2 %>% filter(response_switch == 1)
df3 <- df3 %>% group_by(id) %>%
      mutate(tot_response_switch = sum(response_switch))

df3 <- df3 %>% select(id, id_num, tot_response_switch, accuracy) %>% 
  distinct(id_num, .keep_all = TRUE)

## Visualisation of response switching and accuracy

individual_plot <- ggplot(data = df3, aes(tot_response_switch, accuracy))+geom_line()+geom_point()+ scale_x_continuous(breaks= seq(20, 60, 2))

print(individual_plot)

```

Number of switches between trials did not seem to predict accuracy.

### Survey Response Order

Basically check by selecting relevant columns and going through. Visualisations with number at side and questions at bottom per ppts with reverse code item highlighted. 
#### Autistic Trait Response Order

```{r trait order, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide', eval = FALSE}

## Graphics of response order
## Create list for iteration
id_num  <- c(1:280)

## Make response plot per trail for each participant 
for (i in unique(id_num)){
  
  d <- subset(dfnrc, id_num == i)
  
  d <- d %>%select(id_num, TRAIT_1:TRAIT_12) %>%
  pivot_longer(., cols = c(TRAIT_1:TRAIT_12), names_to = "traits", values_to = "score") 
  
  d <- d %>% mutate(reversed = ifelse(traits == "TRAIT_4", "yes", ifelse(traits == "TRAIT_6", "yes", ifelse(traits == "TRAIT_7", "yes", ifelse(traits == "TRAIT_12", "yes", "no")))))
  
    d <- d %>% mutate(scale = 
                        ifelse(traits ==c("TRAIT_1","TRAIT_4","TRAIT_7", "TRAIT_9","TRAIT_11"), "ss", 
                        ifelse(traits == "TRAIT_3", "c", 
                        ifelse(traits == "TRAIT_6", "c", 
                        ifelse(traits == "TRAIT_8", "c", "as")))))
                                 
  
  order <- factor(d$traits, level = c("TRAIT_1", "TRAIT_2", "TRAIT_3", "TRAIT_4", "TRAIT_5", "TRAIT_6", "TRAIT_7", "TRAIT_8", "TRAIT_9", "TRAIT_10", "TRAIT_11", "TRAIT_12"))
  
plot <- ggplot(data = d, aes(order, score, colour = reversed, group = id_num, shape = scale)) + geom_line(colour = "black")+geom_point(size = 5)+ ggtitle(paste("ID", i)) + scale_y_continuous(limits = c(1, 4)) 

print(plot)
  
}

  
```

The response order for the autistic trait survey was graphically represented. The dataframe used did not include the reversed coded traits so the response order is unchanged. Reverse items and the scale type are depicted within the visualisation. Straight-lining responses and generally straightline responses that did not deviate in terms of reserve coded items was looked for to represent non-compliance. Participant 26 straight-lined to questionnaire only answering with one response. Whereas participants 96, 52, 122 and 138 generally answered within a straight line that did not deviate in terms of reverse questions. 

#### Psychotic-Like Experiences Response Order

```{r ple order, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide', eval = FALSE}

## Graphics of response order
## Create list for iteration
id_num  <- c(1:280)

## Make response plot per trail for each participant 
for (i in unique(id_num)){
  
  d <- subset(df, id_num == i)
  
  d <- d %>%select(id_num, PLE_1:PLE_15) %>%
  pivot_longer(., cols = c(PLE_1:PLE_15), names_to = "ple", values_to = "score") 
  
  
    d <- d %>% mutate(scale = 
                        ifelse(ple == "PLE_1", "pi",
                        ifelse(ple == "PLE_3", "pi",
                        ifelse(ple == "PLE_4", "pi",
                        ifelse(ple == "PLE_5", "pi",   
                        ifelse(ple == "PLE_6", "pi", 
                        ifelse(ple == "PLE_8", "pa", 
                        ifelse(ple == "PLE_10", "pa", 
                        ifelse(ple == "PLE_14", "pa", "be")))))))))
                                 
    order <- factor(d$ple, level = c("PLE_1", "PLE_2", "PLE_3", "PLE_4", "PLE_5", "PLE_6", "PLE_7", "PLE_8", "PLE_9", "PLE_10", "PLE_11", "PLE_12", "PLE_13", "PLE_14", "PLE_15"))
  
 
  
plot <- ggplot(data = d, aes(order, score, colour = scale, group = id_num, shape = scale)) + geom_line(colour = "black")+geom_point(size = 5)+ ggtitle(paste("ID", i)) + scale_y_continuous(limits = c(0, 3)) 

print(plot)
  
}
  
```
 
 The response order for the psychotic-like experiences survey was graphically represented. Scales were colour coded. Four participants responded in complete straight lines with participant 10 and 26 being above 0 and 154 and 166 being at 0. 
 
 

#### Bullying frequency Response Order

```{r bully order, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide', eval = FALSE}

## Graphics of response order
## Create list for iteration
id_num  <- c(1:280)

## Make response plot per trail for each participant 
for (i in unique(id_num)){
  
  d <- subset(df, id_num == i)
  
  d <- d %>%select(id_num, TBV.1_1:TBV.1_8, CBV.1_1:CBV.1_11) %>%
  pivot_longer(., cols = c(TBV.1_1:TBV.1_8, CBV.1_1:CBV.1_11), names_to = "bully", values_to = "score") 
  
  
    d <- d %>% mutate(scale = 
                        ifelse(bully == "TBV.1_1", "tbv",
                        ifelse(bully == "TBV.1_2", "tbv",
                        ifelse(bully == "TBV.1_3", "tbv",
                        ifelse(bully == "TBV.1_4", "tbv",   
                        ifelse(bully == "TBV.1_5", "tbv", 
                        ifelse(bully == "TBV.1_6", "tbv", 
                        ifelse(bully == "TBV.1_7", "tbv", 
                        ifelse(bully == "TBV.1_8", "tbv", "cbv")))))))))
                                 
    order <- factor(d$bully, level = c("TBV.1_1", "TBV.1_2", "TBV.1_3", "TBV.1_4", "TBV.1_5", "TBV.1_6", "TBV.1_7", "TBV.1_8", "CBV.1_1", "CBV.1_2", "CBV.1_3", "CBV.1_4", "CBV.1_5", "CBV.1_6", "CBV.1_7", "CBV.1_8", "CBV.1_9", "CBV.1_10", "CBV.1_11"))
  
 
  
plot <- ggplot(data = d, aes(order, score, colour = scale, group = id_num, shape = scale)) + geom_line(colour = "black")+geom_point(size = 5)+ ggtitle(paste("ID", i)) + scale_y_continuous(limits = c(0, 4)) 

print(plot)
  
}
  
```
 
The response order of both traditional and cyber bullying victimisation scale was graphically visualised. Straight line responses only occurred for 0 representing no bullying experiences. The respondents for these were 2, 5, 8, 16, 17, 22, 37, 46, 48, 79, 81, 90, 108, 116, 161, 166, 170, 188, 190, 201, 215, 228, 239, 243, 246, 267, 271, 272, 278.


### Calculating Questionnaire Indices

#### Autism Trait Index

```{r as index, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

select <- dplyr::select # save as function due to masking

## Create data frame for unique cases
df_as_index <- df %>%  # Calculate index
       group_by(id) %>%
       distinct(id, .keep_all = TRUE) %>%
       mutate(as_index = sum(TRAIT_1,TRAIT_2,TRAIT_3,TRAIT_4,TRAIT_5,
                             TRAIT_6,TRAIT_7, TRAIT_8, TRAIT_9, TRAIT_10,
                             TRAIT_11, TRAIT_12))

## Select relevent variables
df_as_index <- df_as_index %>% select(id, as_index) #remove columns

## Merge datafames
df <- inner_join(df, df_as_index, "id") #merge column


df4 <- df %>% select(id_num, as_index) %>%
             distinct(id_num, .keep_all = TRUE)

## Check for  univariate outliers
boxplot(df4$as_index)
plot(density(df4$as_index))
hist(df4$as_index) 



```
One participant, number 165, scored high on the index

##### Social Skills Subscale

```{r social skills, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Calculate social skill sub scale
## Create data frame for unique cases
df_social_skills <- df %>%  # Social skills scale (Traits 1, 4, 7, 9, 11)
  group_by(id) %>%
  distinct(id, .keep_all = TRUE) %>%
  mutate(social_skills = sum(TRAIT_1,TRAIT_4,TRAIT_7,
                             TRAIT_9, TRAIT_11))

## Select relevant variables
df_social_skills <- df_social_skills %>% select(id, social_skills) #remove columns

## Merge datafames
df <- inner_join(df, df_social_skills, "id") #merge column


df4 <- df %>% select(id_num, social_skills) %>%
             distinct(id_num, .keep_all = TRUE)

## Check for  univariate outliers
boxplot(df4$social_skills)
plot(density(df4$social_skills))
hist(df4$social_skills) 

```

Participant 118, 259, 184 scored low at 5
Participant 248, 165, 18, 219, 160 scored high at 20. 

##### Attention Switching Subscale
```{r attention swtiching, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}
## Calculate attention switching sub scale
## Create data frame for unique cases
df_attention_switching <- df %>%  # Attention switching scale (Traits 2, 5, 10, 12)
  group_by(id) %>%
  distinct(id, .keep_all = TRUE) %>%
  mutate(attention_switching = sum(TRAIT_2,TRAIT_5,TRAIT_10,TRAIT_12))

## Select relevant variables
df_attention_switching <- df_attention_switching %>% select(id, attention_switching) #remove columns

## Merge datafames
df <- inner_join(df, df_attention_switching, "id") #merge column

df4 <- df %>% select(id_num, attention_switching) %>%
             distinct(id_num, .keep_all = TRUE)

## Check for  univariate outliers
boxplot(df4$attention_switching)
plot(density(df4$attention_switching))
hist(df4$attention_switching)

```
Participant 165 is an outlier scoring higher than other participants. 

##### Communication Subscale

```{r attention swtiching, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Calculate communication sub scale
## Create data frame for unique cases
df_communication <- df %>%  # Communication scale (Traits 3, 6, 8)
  group_by(id) %>%
  distinct(id, .keep_all = TRUE) %>%
  mutate(communication = sum(TRAIT_3,TRAIT_6,TRAIT_8))

## Select relevant variables
df_communication <- df_communication %>% select(id, communication) #remove columns

## Merge datafames
df <- inner_join(df, df_communication, "id") #merge column

df4 <- df %>% select(id_num, communication) %>%
             distinct(id_num, .keep_all = TRUE)

## Check for  univariate outliers
boxplot(df4$communication)
plot(density(df4$communication))
hist(df4$communication)

```

No outliers in the communication subscale. 

#### Psychotic-like Experiences Index

```{r ple index, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Calculate PLE sum
## Create data frame for unique cases
df_ple_score <- df %>%  # Calculate score
  group_by(id) %>%
  distinct(id, .keep_all = TRUE) %>%
  mutate(ple_score = sum(PLE_1, PLE_2,PLE_3,PLE_4, PLE_5, PLE_6,
                        PLE_7, PLE_8, PLE_9, PLE_10, PLE_11, PLE_12,
                        PLE_13, PLE_14, PLE_15))

## Select relevant variables
df_ple_score <- df_ple_score %>% select(id, ple_score) #remove columns

## Merge datafames
df <- inner_join(df, df_ple_score, "id") #merge column

df4 <- df %>% select(id_num, ple_score) %>%
             distinct(id_num, .keep_all = TRUE)

## Check for  univariate outliers
boxplot(df4$ple_score)
plot(density(df4$ple_score))
hist(df4$ple_score)

```
Has many outliers above 20. In paticular 185, 3, 156. Other outliers include 264, 65, 213, 153, 39, 194, 15, and 76.

##### Index Without Perceptual Abnormalities

```{r ple_be_pi index, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}
## Calculate ple total without perceptual abnormalities
## Create data frame for unique cases
df_ple_be_pi <- df %>%
  group_by(id) %>%
  distinct(id, .keep_all = TRUE) %>%
  mutate(ple_be_pi = sum(PLE_1, PLE_2, PLE_3, PLE_4, PLE_5, 
                         PLE_6, PLE_7, PLE_9, PLE_11, PLE_12,
                         PLE_13, PLE_15))

## Select relevant variables
df_ple_be_pi <- df_ple_be_pi %>% select(id, ple_be_pi) #remove columns

## Merge datafames
df <- inner_join(df, df_ple_be_pi, "id") #merge column

df4 <- df %>% select(id_num, ple_be_pi) %>%
             distinct(id_num, .keep_all = TRUE)

## Check for  univariate outliers
boxplot(df4$ple_be_pi)
plot(density(df4$ple_be_pi))
hist(df4$ple_be_pi)

```

High outliers again but extreme with only 185, 3 and 156 and 164 being another. 

##### Persecutary Ideation Subscale

```{r pi, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Calculate persecutory ideation sub scale
## Create data frame for unique cases
df_pers_idea <- df %>%  # Persecutory ideation (1, 3, 4, 5, 6)
  group_by(id) %>%
  distinct(id, .keep_all = TRUE) %>%
  mutate(pi_scale = sum(PLE_1, PLE_3,PLE_4,PLE_5, PLE_6))

## Select relevant variables
df_pers_idea <- df_pers_idea %>% select(id, pi_scale) #remove columns

## Merge datafames
df <- inner_join(df, df_pers_idea, "id") #merge column

df4 <- df %>% select(id_num, pi_scale) %>%
             distinct(id_num, .keep_all = TRUE)

## Check for  univariate outliers
boxplot(df4$pi_scale)
plot(density(df4$pi_scale))
hist(df4$pi_scale)

```

Participant 130, 185, 156 and 144 are outliers. 

##### Bizzare Experiences Subscale

```{r be, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Calculate bizzare experiences sub scale
## Create data frame for unique cases
df_bizz_exp <- df %>%  # Bizzare experiences (2, 7, 9, 11, 12, 13,15)
  group_by(id) %>%
  distinct(id, .keep_all = TRUE) %>%
  mutate(be_scale = sum(PLE_2, PLE_7,PLE_9,PLE_11,
                        PLE_12, PLE_13, PLE_15))
                        
## Select relevant variables
df_bizz_exp <- df_bizz_exp %>% select(id, be_scale) #remove columns

## Merge datafames
df <- inner_join(df, df_bizz_exp, "id") #merge column

df4 <- df %>% select(id_num, be_scale) %>%
             distinct(id_num, .keep_all = TRUE)

## Check for  univariate outliers
boxplot(df4$be_scale)
plot(density(df4$be_scale))
hist(df4$be_scale)
```

Outling participants are 185, 3, 156, 213, 264, 65, 194, 39, 229, 76 and 59

##### Perceptual Abnormalities Subscale

```{r pa, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Calculate perceptual abnormalities sub scale
## Create data frame for unique cases
df_perc_abn <- df %>%  # Perceptual abnormalities (8, 10, 14)
  group_by(id) %>%
  distinct(id, .keep_all = TRUE) %>%
  mutate(pa_scale = sum(PLE_8, PLE_10, PLE_14))

## Select relevant variables
df_perc_abn <- df_perc_abn %>% select(id, pa_scale) #remove columns

## Merge datafames
df <- inner_join(df, df_perc_abn, "id") #merge column


df4 <- df %>% select(id_num, pa_scale) %>%
             distinct(id_num, .keep_all = TRUE)

## Check for  univariate outliers
boxplot(df4$pa_scale)
plot(density(df4$pa_scale))
hist(df4$pa_scale)
```

The main outliers are 3, 185, 65, 156

#### Bullying Victimisation Frequency Index
```{r bully index, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Calculate overall bullying victimisation sum
## Create data frame for unique cases
df_bully_score <- df %>% 
  group_by(id) %>%
  distinct(id, .keep_all = TRUE) %>%
  mutate(bully_score = sum(CBV.1_1, CBV.1_2, CBV.1_3, CBV.1_4, CBV.1_5,
                           CBV.1_6, CBV.1_7, CBV.1_8, CBV.1_9, CBV.1_10,
                           CBV.1_11, TBV.1_1, TBV.1_2, TBV.1_3, TBV.1_4,
                           TBV.1_5, TBV.1_6, TBV.1_7, TBV.1_8)) 

## Select relevant variables
df_bully_score <- df_bully_score %>% select(id, bully_score) #remove columns

## Merge datafames
df <- inner_join(df, df_bully_score, "id") #merge column

df4 <- df %>% select(id_num, bully_score) %>%
             distinct(id_num, .keep_all = TRUE)

## Check for  univariate outliers
boxplot(df4$bully_score)
plot(density(df4$bully_score))
hist(df4$bully_score)

```

Outliers include 50, 213, 156, 245, 10, 120, 181, 76, 105, 265, 158, 121, 34, 259, 53.

##### Traditionaly Bullying Victimisation Frequency Subscale
```{r tbv scale, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Calculate traditional bullying victimisation sum
## Create data frame for unique cases
df_tbv_score <- df %>%  
  group_by(id) %>%
  distinct(id, .keep_all = TRUE) %>%
  mutate(tbv_score = sum(TBV.1_1, TBV.1_2, TBV.1_3, TBV.1_4, 
                         TBV.1_5, TBV.1_6, TBV.1_7, TBV.1_8))

## Select relevant variables
df_tbv_score <- df_tbv_score %>% select(id, tbv_score) #remove columns

## Merge datafames
df <- inner_join(df, df_tbv_score, "id") #merge column

df4 <- df %>% select(id_num, tbv_score) %>%
             distinct(id_num, .keep_all = TRUE)

## Check for  univariate outliers
boxplot(df4$tbv_score)
plot(density(df4$tbv_score))
hist(df4$tbv_score)

```

Outliers include 53,  259, 265, 105, 181, 137, 76, 264, 180, 103, 20, 120, 34, 10, 245, 158, 15, 247

##### Cyberbullying Victimisation Frequency Subscale
```{r cbv scale, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Calculate cyber bullying victimisation sum
## Create data frame for unique cases
df_cbv_score <- df %>%  
  group_by(id) %>%
  distinct(id, .keep_all = TRUE) %>%
  mutate(cbv_score = sum(CBV.1_1, CBV.1_2, CBV.1_3, CBV.1_4, CBV.1_5,
                         CBV.1_6, CBV.1_7, CBV.1_8, CBV.1_9, CBV.1_10,
                         CBV.1_11)) 

## Select relevant variables
df_cbv_score <- df_cbv_score %>% select(id, cbv_score) #remove columns

## Merge datafames
df <- inner_join(df, df_cbv_score, "id") #merge column

df4 <- df %>% select(id_num, cbv_score) %>%
             distinct(id_num, .keep_all = TRUE)

## Check for  univariate outliers
boxplot(df4$cbv_score)
plot(density(df4$cbv_score))
hist(df4$cbv_score)

```

The main outliers include 50, 213, 156, 245 and 10

### Trimming Response Time Data
#### Dealing With Extreme Outliers

```{r delete rows, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Delete response time based upon previous_error
df <- df %>% subset(previous_error != "yes")

## Delete rows of data that have errors 
df <- df[!df$correct == "0",]

## Delete rows without switching type information
df <- df[!is.na(df$switch_type), ]

```

df2 <- df %>% select (id_num, switch_type) %>% distinct(switch_type, .keep_all = TRUE) 
Delete data that we are not interested in analysing.

```{r remove extremes, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## See response time per participant
response <- df %>% select(id_num, response_time) %>%
              arrange(response_time)

## Create visualisation of each participants rt per block
individual_plot <- ggplot(data = df, aes(block, response_time, group = id))+geom_line()+geom_point()

print(individual_plot)

## Remove extreme outliers
df <- df %>% filter(response_time < 3000)
df <- df %>% filter(response_time > 200)


## Delete response times 2 SD above and below for each switch type per participant
df <- df %>% group_by(id, switch_type) %>% 
           mutate(avg = mean(response_time), stdev = sd(response_time))

df <- df %>% group_by(id, switch_type) %>%
  filter(response_time <= 2*stdev+avg)
df <- df %>% group_by(id, switch_type) %>%
  filter(response_time >= avg - (2*stdev))

## Create plot
plot(density(df$response_time))
boxplot(df$response_time)

## Look for missing data
df2 <- df %>% select (id_num, switch_type) %>% group_by(switch_type) %>% distinct(id_num, .keep_all = TRUE)

df2 <- df2 %>% filter(switch_type == "global_local")
```

Large reaction time values, above 3000ms, according to the visualisations were removed to not bias trimming of response time data. Trails with response times lower than 200ms were removed as it was deemed unlikely for the to capture the psychological construct due to motor responses and cognitive processing time. Participants response times were then trimmed according to switch type so that systematic bias is not introduced by 2 standard deviations below and above the mean. This was to remove the influence of outliers when creating the response time indices. This procedure left participant 5, 114 and 156 without a global to local switching cost trails.  

### Calculate Switching Response Time Means
#### No Switch Mean
```{r mean no switch, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Calculating mean switching response times per type
df_rt <- df %>% 
  group_by(id, switch_type) %>% 
  summarise(mean_rt = mean(response_time)) 

## Adding no_switch mean column
df_no_switch <- df_rt %>%
  filter(switch_type == "no_switch") #localised no_switch rt

## Rename column
colnames(df_no_switch)[3] <- "no_switch_mean_rt"

## Remove columns
df_no_switch <- df_no_switch %>%
  select(id,no_switch_mean_rt)

## Merge dataframes
df <- inner_join(df, df_no_switch, "id") 

## Ungroup switch typw

df <- df %>% 
  ungroup(switch_type)

df2 <- df %>% select(id_num, no_switch_mean_rt)   %>%
  distinct(id_num, .keep_all = TRUE)

## Create plot
plot(density(df2$no_switch_mean_rt))
boxplot(df2$no_switch_mean_rt)

individual_plot <- ggplot(data = df2, aes(id_num, no_switch_mean_rt, group = id_num))+geom_line()+geom_point()

print(individual_plot)

```
Outliers include 114, 44, 33, 265, 212, 194, 8, 88, 185

#### Global to Local Switch Mean
```{r mean global local switch, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Adding global_local switch mean column
df_global_local <- df_rt %>%
  filter(switch_type == "global_local") #localised global_local rt

## Rename column
colnames(df_global_local)[3] <- "global_local_switch_mean_rt"

## Remove columns
df_global_local <- df_global_local %>%
  select(id,global_local_switch_mean_rt)

## Merge dataframes
df <- inner_join(df, df_global_local, "id") 

## Ungroup switch type
df <- df %>% 
  ungroup(switch_type)

df2 <- df %>%  select(id_num, global_local_switch_mean_rt)   %>%
  distinct(id_num, .keep_all = TRUE)

## Create plot
plot(density(df2$global_local_switch_mean_rt))
boxplot(df2$global_local_switch_mean_rt)

individual_plot <- ggplot(data = df2, aes(id_num, global_local_switch_mean_rt, group = id_num))+geom_line()+geom_point()

print(individual_plot)

```
Outliers include 105, 33, 218, 212, 29, 11. These functions removed participant 5, 114 and 156 from the analysis. 

#### Local to Global Mean
```{r mean local global switch, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Adding local_global switch mean column 
df_local_global <- df_rt %>%
    filter(switch_type == "local_global") #localised local_global rt

## Rename column
colnames(df_local_global)[3] <- "local_global_switch_mean_rt"

## Remove columns
df_local_global <- df_local_global %>%
  select(id,local_global_switch_mean_rt) 

## Merge dataframes
df <- inner_join(df, df_local_global, "id") 

## Ungroup switch type
df <- df %>% 
  ungroup(switch_type)

df2 <- df %>%  select(id_num, local_global_switch_mean_rt)   %>%
  distinct(id_num, .keep_all = TRUE)

## Create plot
plot(density(df2$local_global_switch_mean_rt))
boxplot(df2$local_global_switch_mean_rt)

individual_plot <- ggplot(data = df2, aes(id_num, local_global_switch_mean_rt, group = id_num))+geom_line()+geom_point()

print(individual_plot)

df2 <- df %>%
    group_by(id_num, switch_type) %>%
    summarise(count=n())

```
Some outliers include 19, 265, 212, 29,  194, 11, 238, 33, 69, 205. 

#### Bivariate Means
```{r mean global local no switch, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

df2 <- df %>% select(id_num, no_switch_mean_rt, global_local_switch_mean_rt) %>% distinct(id_num,.keep_all = TRUE)

bagplot(df2$global_local_switch_mean_rt, df2$no_switch_mean_rt)

## Create plot


individual_plot <- ggplot(data = df2, aes(no_switch_mean_rt, global_local_switch_mean_rt))+geom_smooth(method = 'lm') +geom_point()+ 
              scale_y_continuous(limits = c(250, 2500)) + 
              scale_x_continuous(limits = c(250, 2250))

print(individual_plot)

df2 <- df2 %>% filter(id_num != 88)

individual_plot <- ggplot(data = df2, aes(no_switch_mean_rt, global_local_switch_mean_rt))+geom_smooth(method = 'lm') +geom_point()+ 
              scale_y_continuous(limits = c(250, 2500)) + 
              scale_x_continuous(limits = c(250, 2250))

print(individual_plot)

```

Point 88 seems to be a bivariate outlier, remove to check. Does seem to influence relationship.

```{r mean local global no switch, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

df2 <- df %>% select(id_num, no_switch_mean_rt, local_global_switch_mean_rt) %>% distinct(id_num,.keep_all = TRUE)

bagplot(df2$local_global_switch_mean_rt, df2$no_switch_mean_rt)

## Create plot


individual_plot <- ggplot(data = df2, aes(no_switch_mean_rt, local_global_switch_mean_rt))+geom_smooth(method = 'lm') +geom_point()+ 
              scale_y_continuous(limits = c(250, 2000)) + 
              scale_x_continuous(limits = c(250, 2250)) 

print(individual_plot)

df2 <- df2 %>% filter(id_num != 88)

individual_plot <- ggplot(data = df2, aes(no_switch_mean_rt, local_global_switch_mean_rt))+geom_smooth(method = 'lm') +geom_point() + 
              scale_y_continuous(limits = c(250, 2000)) + 
              scale_x_continuous(limits = c(250, 2250))

print(individual_plot)

```

88 seems to be an outlier again. Removing outlier improves fit a little. 


### Calculating Switching Cost Indices

#### Global to Local Switching Cost

```{r switching cost indices, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Global_local switching cost index
## Global-local switching rt minus no switch 
df <- df %>% 
  group_by(id) %>% 
  mutate(global_local_cost_index = global_local_switch_mean_rt - no_switch_mean_rt)

df2 <- df %>% 
  distinct(id, .keep_all = TRUE)

glob_loc_cost<- df2 %>% select(id_num, global_local_cost_index) 

plot(density(df2$global_local_cost_index))
boxplot(df2$global_local_cost_index)

individual_plot <- ggplot(data = df2, aes(id_num, global_local_cost_index, group = id_num))+geom_line()+geom_point()

print(individual_plot)

```


Paticularily 105 being high and 88 being low

#### Local to Global Cost Indices 

```{r local global cost indices, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Local_global switching cost index
## Local-global switching rt minus no-switching rt
df <- df %>% 
  group_by(id) %>% 
  mutate(local_global_cost_index = local_global_switch_mean_rt - no_switch_mean_rt)

df2 <- df %>% 
  distinct(id, .keep_all = TRUE)

loc_glob_cost<- df2 %>% select(id_num, local_global_cost_index) %>%
              arrange(local_global_cost_index)

plot(density(df2$local_global_cost_index))

boxplot(df2$local_global_cost_index)

individual_plot <- ggplot(data = df2, aes(id_num, local_global_cost_index, group = id_num))+geom_line()+geom_point()

print(individual_plot)


```

88, outlier

#### Bivariate Indices
```{r local global cost indices, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

df2 <- df %>% select(id_num, global_local_cost_index, local_global_cost_index) %>% distinct(id_num,.keep_all = TRUE)

bagplot(df2$global_local_cost_index, df2$local_global_cost_index)

## Create plot


individual_plot <- ggplot(data = df2, aes(local_global_cost_index, global_local_cost_index)) +geom_smooth(method = 'lm') + geom_point()

print(individual_plot)

```

### Delete Univariate Outliers in Response Time Data 

```{r remove univariate response outlier, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

df <-df[!(df$id_num == "88" | df$id_num =="105"),] 

```

Participants were removed due to the strong influence biasing results.

### Examining Bivariate Outliers

#### Global to Local Cost Index Outliers

##### Autism Trait Index 

```{r bivarate as trait global local, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

df2 <- df %>% select(id_num, global_local_cost_index, as_index) %>% distinct(id_num,.keep_all = TRUE)

bagplot(cbind(df2$as_index,df2$global_local_cost_index),pch=16,cex=2)

individual_plot <- ggplot(data = df2, aes(as_index, global_local_cost_index)) +geom_smooth(method = 'lm') + geom_point() + 
              scale_y_continuous(limits = c(-400, 600)) + 
              scale_x_continuous(limits = c(10, 50))

print(individual_plot)

## Remove highest outlier
df2 <- df2[!(df2$id_num =="175"),] 


individual_plot <- ggplot(data = df2, aes(as_index, global_local_cost_index))+geom_smooth(method = 'lm') +geom_point() + 
              scale_y_continuous(limits = c(-400, 600)) + 
              scale_x_continuous(limits = c(10, 50))

print(individual_plot)

## Remove rest of outliers
df2 <- df2[!(df2$id_num =="218"| df2$id_num =="25"|
             df2$id_num =="29"),] 


individual_plot <- ggplot(data = df2, aes(as_index, global_local_cost_index))+geom_smooth(method = 'lm') +geom_point() + 
              scale_y_continuous(limits = c(-400, 600)) + 
              scale_x_continuous(limits = c(10, 50))

print(individual_plot)

```

175, 218, 25 and 29 are outliers. 175 strongly influences the relationship direction whereas the other outliers mainly influence the intercept. 175 will be removed from the data later. 

##### Perceptual Abnormalities

```{r bivarate pa scale global local, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

df2 <- df %>% select(id_num, global_local_cost_index, pa_scale) %>% distinct(id_num,.keep_all = TRUE)

bagplot(cbind(df2$pa_scale,df2$global_local_cost_index),pch=16,cex=2)

individual_plot <- ggplot(data = df2, aes(pa_scale, global_local_cost_index)) +geom_smooth(method = 'lm') + geom_point() + 
              scale_y_continuous(limits = c(-400, 600)) + 
              scale_x_continuous(limits = c(0, 10))

print(individual_plot)

## Remove univariate outliers
df2 <- df2[!(df2$id_num =="3" | df2$id_num =="185" |
             df2$id_num =="65"),] 

individual_plot <- ggplot(data = df2, aes(pa_scale, global_local_cost_index)) +geom_smooth(method = 'lm') + geom_point() + 
              scale_y_continuous(limits = c(-400, 600)) + 
              scale_x_continuous(limits = c(0, 10))

print(individual_plot)

```
There are no outliers concerning the bag plot but univariate outliers 3, 185 and 65 are checked to determine if influence direction of relationship. Removing the outliers did not effect the relationship much.  

##### Bizzare Experiences and Persecutary Ideation

```{r bivarate ple be pi global local, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

df2 <- df %>% select(id_num, global_local_cost_index, ple_be_pi) %>% distinct(id_num,.keep_all = TRUE)

bagplot(cbind(df2$ple_be_pi,df2$global_local_cost_index),pch=16,cex=2)

individual_plot <- ggplot(data = df2, aes(ple_be_pi, global_local_cost_index)) +geom_smooth(method = 'lm') + geom_point() + 
              scale_y_continuous(limits = c(-400, 600)) + 
              scale_x_continuous(limits = c(0, 35))


print(individual_plot)

## Check if removed all
df2 <- df2[!(df2$id_num =="175"),] 

individual_plot <- ggplot(data = df2, aes(ple_be_pi, global_local_cost_index)) +geom_smooth(method = 'lm') + geom_point() + 
              scale_y_continuous(limits = c(-400, 600)) + 
              scale_x_continuous(limits = c(0, 35))


print(individual_plot)

## Check if removed all
df2 <- df2[!(df2$id_num =="175"|df2$id_num =="25"),] 

individual_plot <- ggplot(data = df2, aes(ple_be_pi, global_local_cost_index)) +geom_smooth(method = 'lm') + geom_point() + 
              scale_y_continuous(limits = c(-400, 600)) + 
              scale_x_continuous(limits = c(0, 35))


print(individual_plot)

df2 <- df2[!(df2$id_num =="175"|df2$id_num =="25"
             |df2$id_num =="185"|df2$id_num =="3"),] 

individual_plot <- ggplot(data = df2, aes(ple_be_pi, global_local_cost_index)) +geom_smooth(method = 'lm') + geom_point() + 
              scale_y_continuous(limits = c(-400, 600)) + 
              scale_x_continuous(limits = c(0, 35))

print(individual_plot)


df2 <- df2[!(df2$id_num =="175"|df2$id_num =="25"|df2$id_num =="218" 
             |df2$id_num =="185"|df2$id_num =="3"),] 

individual_plot <- ggplot(data = df2, aes(ple_be_pi, global_local_cost_index)) +geom_smooth(method = 'lm') + geom_point() + 
              scale_y_continuous(limits = c(-400, 600)) + 
              scale_x_continuous(limits = c(0, 35))

print(individual_plot)


```
The same outliers that influenced autism trait index is in this 175, 218 and 25. In addition to 185 and 3 as univariate outliers in perceptual abnormalities. As outliers have an impact on the relationship and removing them will make other outliers stronger in other relationships. All outliers will be removed.


##### Bullying 

```{r bivarate pa scale global local, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

df2 <- df %>% select(id_num, bully_score, global_local_cost_index) %>% distinct(id_num,.keep_all = TRUE)

bagplot(cbind(df2$buly_score,df2$global_local_cost_index))

individual_plot <- ggplot(data = df2, aes(bully_score, global_local_cost_index)) +geom_smooth(method = 'lm') + geom_point()

print(individual_plot)

```

Again, 175, 218 and 25 are outliers

#### Local to Global Cost Index Outliers

##### Autism Trait Index

```{r bivarate as trait local global, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

df2 <- df %>% select(id_num, local_global_cost_index, as_index) %>% distinct(id_num,.keep_all = TRUE)

bagplot(cbind(df2$as_index,df2$local_global_cost_index),pch=16,cex=2)

individual_plot <- ggplot(data = df2, aes(as_index, local_global_cost_index)) +geom_smooth(method = 'lm') + geom_point() + 
              scale_x_continuous(limits = c(10, 50)) + 
              scale_y_continuous(limits = c(-300, 600))

print(individual_plot)

df2 <- df2[!(df2$id_num =="259"|df2$id_num =="165"),] 

individual_plot <- ggplot(data = df2, aes(as_index, local_global_cost_index)) +geom_smooth(method = 'lm') + geom_point()+ scale_x_continuous(limits = c(10, 50)) + 
              scale_y_continuous(limits = c(-300, 600))

print(individual_plot)

df2 <- df2[!(df2$id_num =="164"|df2$id_num =="19"),] 

individual_plot <- ggplot(data = df2, aes(as_index, local_global_cost_index)) +geom_smooth(method = 'lm') + geom_point()+ scale_x_continuous(limits = c(10, 50)) + 
              scale_y_continuous(limits = c(-300, 600))

print(individual_plot)

```

Outliers most likely to effect relationship is 259 and 165. Removing them does not effect the relationship much. 164 and 19 hardly effect relationship. 

##### Perceptual Abnormalities

```{r bivarate pa scale local global, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

df2 <- df %>% select(id_num, local_global_cost_index, pa_scale) %>% distinct(id_num,.keep_all = TRUE)

bagplot(cbind(df2$pa_scale,df2$local_global_cost_index),pch=16,cex=2)

individual_plot <- ggplot(data = df2, aes(pa_scale, local_global_cost_index)) +geom_smooth(method = 'lm') + geom_point() + 
              scale_y_continuous(limits = c(-300, 600)) + 
              scale_x_continuous(limits = c(0, 10))

print(individual_plot)

df2 <- df2[!(df2$id_num =="3"|df2$id_num =="185"|df2$id_num =="65"),]

individual_plot <- ggplot(data = df2, aes(pa_scale, local_global_cost_index)) +geom_smooth(method = 'lm') + geom_point() + 
              scale_y_continuous(limits = c(-300, 600)) + 
              scale_x_continuous(limits = c(0, 10))

print(individual_plot)

``` 

The univariate outliers stronger influences the relationship, these will be removed due to previous analyses anyway. 

##### Bizzare Experiences and Persecutary Ideation

```{r bivarate ple be pi local global, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

df2 <- df %>% select(id_num, local_global_cost_index, ple_be_pi) %>% distinct(id_num,.keep_all = TRUE)

bagplot(cbind(df2$ple_be_pi,df2$local_global_cost_index),pch=16,cex=2)

individual_plot <- ggplot(data = df2, aes(ple_be_pi, local_global_cost_index)) +geom_smooth(method = 'lm') + geom_point() + 
              scale_y_continuous(limits = c(-300, 600)) + 
              scale_x_continuous(limits = c(0, 35))


print(individual_plot)

df2 <- df2[!(df2$id_num =="3"|df2$id_num =="185"),]
individual_plot <- ggplot(data = df2, aes(ple_be_pi, local_global_cost_index)) +geom_smooth(method = 'lm') + geom_point() + 
              scale_y_continuous(limits = c(-300, 600)) + 
              scale_x_continuous(limits = c(0, 35))


print(individual_plot)


df2 <- df2[!(df2$id_num =="3"|df2$id_num =="164"),]
individual_plot <- ggplot(data = df2, aes(ple_be_pi, local_global_cost_index)) +geom_smooth(method = 'lm') + geom_point() + 
              scale_y_continuous(limits = c(-300, 600)) + 
              scale_x_continuous(limits = c(0, 35))


print(individual_plot)

```
185 and 3 influence the relationship but this will be removed anyway. 164 slightly influences the relationship


##### Bullying 

```{r bivarate bully local global, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

df2 <- df %>% select(id_num, bully_score, local_global_cost_index) %>% distinct(id_num,.keep_all = TRUE)

bagplot(cbind(df2$buly_score,df2$local_global_cost_index))

individual_plot <- ggplot(data = df2, aes(bully_score, local_global_cost_index)) +geom_smooth(method = 'lm') + geom_point()

print(individual_plot)

```
No outliers


#### Remove Reaction Time Outliers 
```{r bivarate bully local global, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

df <- df[!(df$id_num =="175"|df$id_num =="218"| 
           df$id_num =="25"|df$id_num =="29"| 
           df$id_num =="3"|df$id_num =="185"| 
           df$id_num =="65"|df$id_num =="164"| 
           df$id_num =="19"|df$id_num =="165"| 
           df$id_num =="259"),]

```

Due to the influence of univariate and bivariate outliers influence on many relationships. All are removed from analysis.

#### Bizzare Experiences and Persecutary Ideation Outliers

##### Autism Trait Index

```{r bivarate as trait local global, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

df2 <- df %>% select(id_num, ple_be_pi, as_index) %>% distinct(id_num,.keep_all = TRUE)

bagplot(cbind(df2$as_index,df2$ple_be_pi),pch=16,cex=2)

individual_plot <- ggplot(data = df2, aes(as_index, ple_be_pi)) +geom_smooth(method = 'lm') + geom_point() + 
              scale_x_continuous(limits = c(10, 50)) + 
              scale_y_continuous(limits = c(0, 30))

print(individual_plot)
```
No outliers.


##### Perceptual Abnormalities

```{r bivarate pa scale local global, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

df2 <- df %>% select(id_num, ple_be_pi, pa_scale) %>% distinct(id_num,.keep_all = TRUE)

bagplot(cbind(df2$pa_scale,df2$ple_be_pi),pch=16,cex=2)

individual_plot <- ggplot(data = df2, aes(pa_scale, ple_be_pi)) +geom_smooth(method = 'lm') + geom_point() 

print(individual_plot)

```

No outliers. 

##### Bullying 

```{r bully ple_be_pi bivariate, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

df2 <- df %>% select(id_num, bully_score, ple_be_pi) %>% distinct(id_num,.keep_all = TRUE)

bagplot(cbind(df2$buly_score,df2$ple_be_pi))

individual_plot <- ggplot(data = df2, aes(bully_score, ple_be_pi)) +geom_smooth(method = 'lm') + geom_point()

print(individual_plot)

```

No outliers.

## Switching Model

### Assumption: Absence of Multicolinearity

```{r multicolinearlity, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

df2 <- df %>% distinct(id_num, .keep_all = TRUE)

## Check correlation of variables
cor(df2[,c(62,67,70,79,80)], use="complete.obs") # No high correlations

```


### Determining Distribution Fit
```{r distributions, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

df2 <- df %>% 
  distinct(id, .keep_all = TRUE)

## Temp load package
library(fitdistrplus)


## Show empirical density and cumulative distribution
plotdist(df2$local_global_cost_index, histo = TRUE, demp = TRUE)
plotdist(df2$global_local_cost_index, histo = TRUE, demp = TRUE)

## Skewness-kurtosis plot 
descdist(df2$local_global_cost_index, boot = 1000, discrete = FALSE)
descdist(df2$global_local_cost_index, boot = 1000, discrete = FALSE)

### Transform to positive values to determine best fit
## adding 1000 to each score
add <- function(x){
  #add to reflect
  1000 + x
}

df2$local_global_cost_index <- sapply(df2$local_global_cost_index, add)
df2$global_local_cost_index <- sapply(df2$global_local_cost_index, add)



## Fit distributions

fit_norm <- fitdist(df2$local_global_cost_index, "norm")
plot(fit_norm)

fit_lnorm <- fitdist(df2$local_global_cost_index, "lnorm")
plot(fit_lnorm)

fit_gamma <- fitdist(df2$local_global_cost_index, "gamma")
plot(fit_gamma)

fit_logis <- fitdist(df2$local_global_cost_index, "logis")
plot(fit_logis) # Appears to follow logistic distribution well


fit_norm1 <- fitdist(df2$global_local_cost_index, "norm")
plot(fit_norm1)

fit_lnorm1 <- fitdist(df2$global_local_cost_index, "lnorm")
plot(fit_lnorm1)

fit_gamma1 <- fitdist(df2$global_local_cost_index, "gamma")
plot(fit_gamma1)

fit_logis1 <- fitdist(df2$global_local_cost_index, "logis")
plot(fit_logis1) # Appears to follow logistic distribution well




## Look at AIC's
fit_norm$aic
fit_lnorm$aic
fit_gamma$aic
fit_logis$aic # Best fit


fit_norm1$aic 
fit_lnorm1$aic
fit_gamma1$aic
fit_logis1$aic # Best fit

detach(package:fitdistrplus)


```

Lognormal best fit for analysis.

#### Local to Global Model 
```{r comparisons, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Select variables 
df2 <- df2 %>% 
  distinct(id, .keep_all = TRUE) %>% 
  select(id, id_num, local_global_cost_index, global_local_cost_index, pa_scale, as_index, ple_be_pi, be_scale, pi_scale, AGE)


## Log Transformed
lmod_locswitch_log <- lm(log(local_global_cost_index) ~  as_index + pa_scale, data = df2)
summary(lmod_locswitch_log) 

## Normal
lmod_locswitch <- lm(local_global_cost_index ~  pa_scale + as_index, data = df2)
summary(lmod_locswitch) 


```



#### Homoscadasticity of Residuals
```{r glm residual normality, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}




## Checking homoscadescity of residual
residualPlot(lmod_locswitch, type = 'rstudent')

bptest(lmod_locswitch) #  sig different to heteroscatic variance (not homoscadestic)


```
#### Global to Local Model Comparisons
##### Normal distrbution 
```{r comparisons, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}




lmod_gloswitch <- lm(global_local_cost_index ~  as_index + pa_scale, data = df2)
summary(lmod_gloswitch)


## Checking homoscadescity of residual
residualPlot(lmod_gloswitch, type = 'rstudent')

bptest(lmod_gloswitch) #  sig different to heteroscatic variance (not homoscadestic)




```


##### linear model
```{r comparisons, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}


lm_gloswitch <- lm(global_local_cost_index ~ as_index, data = df2) 
summary(lm_gloswitch) # Not significant rs = 0.004

lm_gloswitch1 <- lm(global_local_cost_index ~  ple_be_pi + as_index, data = df2)
summary(lm_gloswitch1) ## slightly improves fit to 0.005

lm_gloswitch2 <- lm(global_local_cost_index ~  ple_be_pi + as_index + pa_scale, data = df2)
summary(lm_gloswitch2) ## reduces rs = 0.0016

lm_gloswitch3 <- lm(global_local_cost_index ~  as_index + pa_scale, data = df2)
summary(lm_gloswitch3) ## without ple_be_pi rs = 0.003

lm_gloswitch4 <- lm(global_local_cost_index ~  as_index*pa_scale + ple_be_pi, data = df2)
summary(lm_gloswitch4) ## Moderating effect shows positive relationship


```


#### Normality of Residuals and Homoscadasticity 
```{r glm residual normality, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}


## Check normality of residuals
plot(density(resid(lm_gloswitch2)))

hist(resid(lm_gloswitch2)) 

qqPlot(resid(lm_gloswitch2))

shapiro.test(resid(lm_gloswitch2)) 

## Check fit of model
plot(fitted(lm_gloswitch2),df2$global_local_cost_index,)
abline(a = 0, b = 1)


## Checking homoscadescity of residual
residualPlot(lm_gloswitch2, type = c('deviance'))

bptest(lm_gloswitch2) #  sig different to heteroscatic variance (not homoscadestic)



library(gamlss)
## See summary of model fit
plot(lm_gloswitch2)

## Check Worm Plot
wp(lm_gloswitch2, ylim.all = 1, xlim.all = 3) # Lower value looks like outlier


detach(package:gamlss)

```

## Bullying Model

### Multicolinearlity 


### Determining Distribution Fit
```{r distributions, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

df2 <- df %>% 
  distinct(id, .keep_all = TRUE)

## Temp load package
library(fitdistrplus)


## Show empirical density and cumulative distribution
plotdist(df2$bully_score, histo = TRUE, demp = TRUE)


## Skewness-kurtosis plot 
descdist(df2$bully_score, boot = 1000, discrete = FALSE)


## Fit distribution

fit_norm <- fitdist(df2$bully_score, "norm")
plot(fit_norm)

fit_exp <- fitdist(df2$bully_score, "exp")
plot(fit_exp)


## Look at AIC's
fit_norm$aic
fit_exp$aic

detach(package:fitdistrplus)


df2 <-df %>% distinct(id, .keep_all = TRUE)
df2$AGE <- as.numeric(df2$AGE)
summary(df2)
sd(df2$AGE)


```


### Model Comparison
##### Exponential Distrbution GLM
```{r comparisons, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}


## Select variables 
df2 <- df2 %>% 
  distinct(id, .keep_all = TRUE) %>% 
  select(id, id_num, local_global_cost_index, global_local_cost_index, pa_scale, as_index, ple_be_pi, be_scale, pi_scale, AGE, bully_score)


### Transform to positive values to determine best fit
## adding 1 to each score
add <- function(x){
  #add to reflect
  1 + x
}

df2$bully_score <- sapply(df2$bully_score, add)

lmod_bully3 <- lm(log(bully_score) ~  as_index + pa_scale,  data = df2)
summary(lmod_bully3)



## Load package
library(gamlss)


lmod_bully3 <- gamlss(bully_score ~  as_index + pa_scale, family = EXP(mu.link = "identity"),  data = df2)
summary(lmod_bully3) ## Both as and pa significant together AIC = 1710



Rsq(lmod_bully3, type = c("both"))

detach(package:gamlss)

confint(lmod_bully3)


```

##### Normality and Homoscadasticity of Residuals



```{r comparisons, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}
## Check normality of residuals
plot(density(resid(lmod_bully3)))

hist(resid(lmod_bully3)) 

qqPlot(resid(lmod_bully3))

shapiro.test(resid(lmod_bully3)) 

## Check fit of model
plot(fitted(lmod_bully3),df2$bully_score,)
abline(a = 0, b = 1)


## Checking homoscadescity of residual
residualPlot(lmod_bully3, type = c('weighted'))

bptest(lmod_bully3) #  sig different to heteroscatic variance (not homoscadestic)



library(gamlss)
## See summary of model fit
plot(lmod_bully3)

## Check Worm Plot
wp(lmod_bully3, ylim.all = 1, xlim.all = 3) # Lower value looks like outlier

## Check term plot
term.plot(lmod_bully3)

detach(package:gamlss)

```


## Bizzare Exerpiences and Persecutary Ideation Model

### Multicolinearity

### Determining Distribution Fit
```{r distributions, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

df2 <- df %>% 
  distinct(id, .keep_all = TRUE)

### Transform to positive values to determine best fit
## adding 1 to each score
add <- function(x){
  #add to reflect
  1 + x
}



df2$ple_be_pi <- sapply(df2$ple_be_pi, add)


## Temp load package
library(fitdistrplus)


## Show empirical density and cumulative distribution
plotdist(df2$ple_be_pi, histo = TRUE, demp = TRUE)

## Skewness-kurtosis plot 
descdist(df2$ple_be_pi, boot = 1000, discrete = FALSE)



## Fit distribution

fit_norm <- fitdist(df2$ple_be_pi, "norm")
plot(fit_norm)

fit_lnorm <- fitdist(df2$ple_be_pi, "lnorm")
plot(fit_lnorm)

fit_gamma <- fitdist(df2$ple_be_pi, "gamma")
plot(fit_gamma)

fit_logis <- fitdist(df2$ple_be_pi, "logis")
plot(fit_logis)

fit_weibull <- fitdist(df2$ple_be_pi, "weibull")
plot(fit_weibull)


## Look at AIC's
fit_norm$aic
fit_lnorm$aic
fit_gamma$aic
fit_logis$aic
fit_weibull$aic

detach(package:fitdistrplus)


```
Although states log norm is best fit QQ plot appears otherwise. Gamma looks more appears a better choice.

### Model Comparison
##### Gamma Distrbution GLM
```{r comparisons, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}


## Select variables 
df2 <- df2 %>% 
  distinct(id, .keep_all = TRUE) %>% 
  select(id, id_num, local_global_cost_index, global_local_cost_index, pa_scale, as_index, ple_be_pi, be_scale, pi_scale, AGE, bully_score)



lmod_ple2 <- lm(log(ple_be_pi) ~  as_index + pa_scale , data = df2)
summary(lmod_ple2)



residualPlot(lmod_ple2, type = "deviance")

bptest(lmod_ple2) 



## Load package
library(gamlss)






## Building Models
lmod_ple <- gamlss(ple_be_pi ~  as_index , family = GA, data = df2)
summary(lmod_ple) ## Significant (0.002) AIC = 1707 SBC = 1723

lmod_ple2 <- gamlss(ple_be_pi ~  as_index + pa_scale , family = GA, data = df2)
summary(lmod_ple2) ## Both predict independently AIC = 1700

lmod_ple3 <- gamlss(ple_be_pi ~  as_index + pa_scale + bully_score , family = GA, data = df2)
summary(lmod_ple3) ## Bully score removes other significance AIC = 1300

lmod_ple4 <- gamlss(ple_be_pi ~  as_index + bully_score , family = GA, data = df2)
summary(lmod_ple4) ## Accounts for bullying even alone AIC 1298

lmod_ple5 <- gamlss(ple_be_pi ~  as_index*bully_score + pa_scale , family = GA, data = df2)
summary(lmod_ple5) ## 1277 AIC

lmod_ple6 <- gamlss(ple_be_pi ~  as_index*bully_score, family = GA, data = df2)
summary(lmod_ple6) ## 1275

```



##### Lognorm GLM
```{r comparisons, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}


lmod_ple <- gamlss(ple_be_pi ~  as_index , family = LOGNO, data = df2)
summary(lmod_ple) ## Significant (0.002) AIC = 1707 SBC = 1723

lmod_ple2 <- gamlss(ple_be_pi ~  as_index + pa_scale , family = LOGNO, data = df2)
summary(lmod_ple2) ## Both predict independently AIC = 1700

lmod_ple3 <- gamlss(ple_be_pi ~  as_index + pa_scale + bully_score , family = LOGNO, data = df2)
summary(lmod_ple3) ## Bully score removes other significance AIC = 1300

lmod_ple4 <- gamlss(ple_be_pi ~  as_index + bully_score , family = LOGNO, data = df2)
summary(lmod_ple4) ## Accounts for bullying even alone AIC 1298

lmod_ple5 <- gamlss(ple_be_pi ~  as_index*bully_score + pa_scale , family = LOGNO, data = df2)
summary(lmod_ple5) ## 1277 AIC

lmod_ple6 <- gamlss(ple_be_pi ~  as_index*bully_score, family = LOGNO, data = df2)
summary(lmod_ple6) ## 1275


detach(package:gamlss)


```

Get similar results with log normal transformation anyway. 

##### Normality and Homoscadasticity of Residuals








## SEM

```{r comparisons, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

df2 <- df %>% 
  distinct(id, .keep_all = TRUE) 

summary(df2[, c('be_scale', 'pi_scale', 'cbv_score', 'tbv_score', 'communication', 'social_skills', 'attention_switching', 'pa_scale')])

library(MVN)
test <- mvn(df2[, c('be_scale', 'pi_scale', 'cbv_score', 'tbv_score', 'communication', 'social_skills', 'attention_switching', 'pa_scale')], mvnTest = 'royston')

test$multivariateNormality # Not multivariate normality

## Main analysis check

lmod_sem <- lm(log(ple_be_pi) ~  as_index + pa_scale + bully_score , data = df2)
summary(lmod_sem)

bptest(lmod_sem) 

lmod_sem2 <- lm(log(bully_score) ~  as_index + pa_scale,  data = df2)
summary(lmod_sem2)
bptest(lmod_sem2) 


## Change labels for visualisation 
colnames(df2)[70] <- "BE"
colnames(df2)[69] <- "PI"
colnames(df2)[71] <- "PA"
colnames(df2)[73] <- "TBV"
colnames(df2)[74] <- "CBV"
colnames(df2)[66] <- "C"
colnames(df2)[65] <- "AS"
colnames(df2)[64] <- "SS"

library(lavaan)



sem <- '
## CFA model
DLE =~ NA*BE + PI
BV =~ NA*TBV + CBV
AT =~ NA*C + SS + AS
DLE ~~ 1*DLE
AT ~~ 1*AT
BV ~~ 1*BV

## Path model
DLE ~ b*BV + c*AT + d*PA 
BV ~ a*AT + e*PA

dir1 := c
ind1 := a*b
dir2 := d
ind2 := e*b
dir3 := b
total := dir1 + ind1 + dir2 + ind2 + dir3

'
fit <- sem(sem, data = df2, se = 'robust.sem', test = 'satorra.bentler')
summary(fit, fit.measures = TRUE, standardized=TRUE)


df3 <- df2 %>% filter(GND != "Non-binary / third gender")

fit_con <- sem(sem, data = df3, group = "GND",  se = 'robust.sem', test = 'satorra.bentler')
summary(fit_con, fit.measures = TRUE, standardized=TRUE)

fit_met <- sem(sem, data = df3, group = "GND",  group.equal = 'loadings',se = 'robust.sem', test = 'satorra.bentler')
summary(fit_met, fit.measures = TRUE, standardized=TRUE)

library(semTools)

summary(compareFit(fit_con, fit_met))

boot <- parameterestimates(fit, level = 0.95, boot.ci.type ='bca.simple', standardized = T)

print(boot)

summary(boot)
boot2 <-bootstrapLavaan(fit, R=1000)
summary(boot)



library(semTools)

summary(compareFit(fit, fit2))


modindices(fit)
detach(package:lavaan)



```


## Visualisation 

```{r comparisons, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

library(lavaan)
library(tidySEM)
library(dplyr)

get_nodes(fit)

lay <- get_layout("","","pa_scale","","","","","", 
                  "","","","","bully","","ple", "", 
                  "","", "at","", "", "","", "",
                  "social_skills","communication","attention_switching", "","tbv_score","cbv_score","be_scale","pi_scale", rows = 4)

graph_sem(fit, layout = lay)

```



```{r comparisons, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

library(lavaan)
library(semPlot)
library(gplots)




pathdiagram = semPaths(fit, layout = "tree",what="std",sizeLat=10, 
                       sizeLat2=10,sizeMan=6, sizeMan2=6,fade= F,
                       edge.color = "#000000", 
                       edge.label.cex=1.2, edge.label.position=.5,
                       asize=3,rotation=3,mar=c(6,6,6,12), 
                       whatLabels = "est", optimizeLatRes = T, esize = 10,
                       levels = c(1.5,3,4.5,5.9))

edgelabels = pathdiagram$graphAttributes$Edges$labels  #this and the next few lines remove leading zeros from factor loading coefficients for easier reading
edgelabels[which(pathdiagram$graphAttributes$Edges$curve == 0)] = gsub("0[.]", ".", edgelabels[which(pathdiagram$graphAttributes$Edges$curve == 0)])
pathdiagram$graphAttributes$Edges$labels = edgelabels




# we will also make the structural relationships in the mediation model heavier (thicker lines) so that part is easier to see
latentnodes = which(pathdiagram$Arguments$shape == "circle")  # identify which nodes correspond with latent variables
latentedges = which((pathdiagram$Edgelist$from %in% latentnodes) & (pathdiagram$Edgelist$to %in% latentnodes))  # identify which nodes are connecting latent variables
pathdiagram$graphAttributes$Edges$width[latentedges] = 5
pathdiagram$graphAttributes$Edges$label.font[latentedges] = 2
pathdiagram$graphAttributes$Edges$fade[latentedges] = 10

# we will also make the structural relationships in the mediation model heavier (thicker lines) so that part is easier to see
panode = which(pathdiagram$Arguments$labels == "PA")  # identify which nodes correspond with latent variables 
paedge = which(pathdiagram$Edgelist$from %in% panode)  # identify which nodes are connecting latent variables
pathdiagram$graphAttributes$Edges$width[paedge] = 5
pathdiagram$graphAttributes$Edges$label.font[paedge] = 2
pathdiagram$graphAttributes$Edges$fade[paedge] = 10
pathdiagram$graphAttributes$Nodes$width[panode] = 10
pathdiagram$graphAttributes$Nodes$height[panode] = 10

latent1 <- which(pathdiagram$Arguments$labels == c("BE","PI"))
latent1edge = which(pathdiagram$Edgelist$to %in% latent1)
pathdiagram$graphAttributes$Edges$color[latent1edge] = "#2B728A"

latent5 <- which(pathdiagram$Arguments$labels == "CBV")
latent5edge = which(pathdiagram$Edgelist$to %in% latent5)
pathdiagram$graphAttributes$Edges$color[latent5edge] = "#04347D"

latent6 <- which(pathdiagram$Arguments$labels == "TBV")
latent6edge = which(pathdiagram$Edgelist$to %in% latent6)
pathdiagram$graphAttributes$Edges$color[latent6edge] = "#04347D"

latent3 <- which(pathdiagram$Arguments$labels == c( "AS", "C", "SS"))
latent3edge = which(pathdiagram$Edgelist$to %in% latent3)
pathdiagram$graphAttributes$Edges$color[latent3edge] = "#46085E"



# add fit statistics to the path diagram image
fitstats = fitMeasures(fit, c("chisq.scaled","df","pvalue.scaled","cfi.robust","tli.robust","rmsea.robust"))


plot(pathdiagram)

text("(a)", x=-0.25, y= -.05, font=3, srt=0, cex=0.9, adj=0.5)
text("(c)", x=0.25, y= -.05, font=3, srt=0, cex=0.9, adj=0.5)
text("(b)", x= 0, y= .45, font=3, srt=0, cex=0.9, adj=0.5)
text("(e)", x= -0.25, y= .73, font=3, srt=0, cex=0.9, adj=0.5)
text("(d)", x= 0.25, y= .73, font=3, srt=0, cex=0.9, adj=0.5)

text("***", x= 0.06, y= .36, font=3, srt=0, cex=0.9, adj=0.5)
text("***",x= 0.24, y= .68, font=3, srt=0, cex=0.9, adj=0.5)
text("**", x= -0.12, y= .01, font=3, srt=0, cex=0.9, adj=0.5)
text("**",x= -0.13, y= .68, font=3, srt=0, cex=0.9, adj=0.5)

fitstring1 = paste0("Fit Indices:")
fitstring2 = paste0("(",fitstats["df"],") = ",sprintf("%.2f",fitstats["chisq.scaled"]))
fitstring3 = paste0("p = ",sprintf("%.2f",fitstats["pvalue.scaled"]))
fitstring4 = paste0("Robust CFI = ", sprintf("%.2f",fitstats["cfi.robust"]))
fitstring5 = paste0("Robust TLI = ",sprintf("%.2f",fitstats["tli.robust"]))
fitstring6 = paste0("Robust RMSEA = ",sprintf("%.2f",fitstats["rmsea.robust"]))
fitstring7 = paste0("")
fitstring8 = paste0("Specific Indirect Effects:")
fitstring9 = paste0("a*b =  0.0807*")
fitstring10 = paste0("[CI = 0.0072, 0.1541] (5.8%)")
fitstring11 = paste0("e*b =  0.1412*")
fitstring12 = paste0("[CI = 0.0072, 0.1541] (9.1%)")
fitstringfull = paste(fitstring1, fitstring2, fitstring3, fitstring4, fitstring5,fitstring6, fitstring7,fitstring8, fitstring9,fitstring10, fitstring11, fitstring12,sep="\n")
text(2, 0.7, fitstringfull, adj=c(1,1), family="optima", cex=1.2, xpd=NA)

dev.copy(png,'mediation.png', res=2000, width=300, height=200, unit="mm")
dev.off()

```
Line thickness weighted for standardised parameter estimates (not for structural model but for measurement model)


## Scatterplots with latent variable estimates

```{r comparisons, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}


## Extract coefecients from SEM
a = coef(fit)["a"]
b = coef(fit)["b"]
c = coef(fit)["c"]
d = coef(fit)["d"]
e = coef(fit)["e"]
ab = a*b
eb = e*b

 


## Construct SEM without mediation to get total effect

sem2 <- '
## CFA model
DLE =~ NA*BE + PI
BV =~ NA*TBV + CBV
AT =~ NA*C + SS + AS
DLE ~~ 1*DLE
AT ~~ 1*AT
BV ~~ 1*BV

## Path model
DLE ~ c*AT 

'
fittotal <- sem(sem2, data = df2, se = 'robust.sem', test = 'satorra.bentler',std.lv=T )
summary(fittotal, fit.measures = TRUE, standardized=TRUE)

ctotal = coef(fittotal)["c"]


sem3 <- '
## CFA model
DLE =~ NA*BE + PI
BV =~ NA*TBV + CBV
AT =~ NA*C + SS + AS
DLE ~~ 1*DLE
AT ~~ 1*AT
BV ~~ 1*BV

## Path model
DLE ~ d*PA

'
fittotal2 <- sem(sem3, data = df2, se = 'robust.sem', test = 'satorra.bentler',std.lv=T )
summary(fittotal2, fit.measures = TRUE, standardized=TRUE)
dtotal = coef(fittotal2)["d"]



test1 = coef(fittotal2)["d"] # BV - increases from  .783 to 0.861

test2 = coef(fittotal2)["d"] # AT - increases from .783 to 0.815

test3 = coef(fittotal2)["d"] # AT + BV - increases from .783 to 0.867

## Use CFA to obtain latent estimates


cfa <- '
## CFA model
DLE =~ NA*BE + PI
BV =~ NA*TBV + CBV
AT =~ NA*C + SS + AS
DLE ~~ 1*DLE
AT ~~ 1*AT
BV ~~ 1*BV

'
fitcfa <- cfa(cfa, data=df2, std.lv=T)
latent_estimates <- data.frame(lavPredict(fitcfa))

# Plot 1 AT and BV (path a)
png(paste0("Mediation_raw_path_a.png"), width=75, height=75, unit="mm", res=1000, pointsize=8)   
plot(y=latent_estimates$BV, x=latent_estimates$AT, ylab="Latent Bullying-Victimisation (M) Estimate", xlab="Latent Autistic Traits (X) Estimate", main="Autistic traits to Bullying-Victimiation (Path a)", pch=16, ps=1.25, cex.lab=1, cex.axis=1, cex.main=1)
abline(a=0, b= a)   # draw regression line
text(2.9, 2.5, "a", font=3, cex=1.25)   # add "a" symbol clarifying what the regression line is
dev.off()


# Plot 2 BV and DLE (path b)
png(paste0("Mediation_raw_path_b.png"), width=75, height=75, unit="mm", res=1000, pointsize=8)   
plot(y=latent_estimates$DLE, x=latent_estimates$BV, ylab="Latent Delusional-Like Experiences (Y) Estimate", xlab="Latent Bullying-Victimisation (M) Estimate", main="Bullying-Victimisation to Delusional-Like Experiences (path b)", pch=16, ps=1.25, cex.lab=1, cex.axis=1, cex.main=1)
abline(a=0, b= b)  
text(2.9, 1.1, "b", font=3, cex=1.25)
dev.off()

# plot 3 AT and DLE (path c)
png(paste0("Mediation_raw_path_c.png"), width=75, height=75, unit="mm", res=1000, pointsize=8)   
plot(y=latent_estimates$DLE, x=latent_estimates$AT, ylab="Delusional-Like Experiences (Y) Estimate", xlab="Autistic Traits (X) Estimate", main="Autistic Traits to Delusional-Like Experiences (path c)", pch=16, ps=1.25, cex.lab=1, cex.axis=1, cex.main=1)
abline(a=0, b= ctotal, lty=2)   # plot c (total effect) regression line
abline(a=0, b= c, lty=1)   # plot c' (direct effect) regression line
text(2.9, 2.7, "c", font=3, cex=3)
text(2.9, 1.4, "c'", font=3, cex=3)
#legend("topleft", lty=c(1,2), legend=c("c (total effect)", "c' (direct effect)"), bty="n")
dev.off()

# plot 4 PA to BV (path e)
png(paste0("Mediation_raw_path_e.png"), width=75, height=75, unit="mm", res=1000, pointsize=8)   
plot(y=latent_estimates$BV, x= df2$PA, ylab="Latent Bullying-Victimisation (M) Estimate", xlab="Perceptual Abnormalities Score", main="Perceptual Abnormalities to Bullying-Victimiation (Path e)", pch=16, ps=1.25, cex.lab=1, cex.axis=1, cex.main=1)
abline(a=0, b= e)   # draw regression line
text(2.9, 2.5, "e", font=3, cex=1.25)   # add "a" symbol clarifying what the regression line is
dev.off()


# Plot 5 PA and DLE (path d)
png(paste0("Mediation_raw_path_d.png"), width=75, height=75, unit="mm", res=1000, pointsize=8)   
plot(y=latent_estimates$DLE, x=df2$PA, ylab="Delusional-Like Experiences (Y) Estimate", xlab="Perceptual Abnormalities Score", main="Perceptual Abnormalities to Delusional-Like Experiences (path d)", pch=16, ps=1.25, cex.lab=1, cex.axis=1, cex.main=1)
abline(a=0, b= dtotal, lty=2)   # plot c (total effect) regression line
abline(a=0, b= d, lty=1)   # plot c' (direct effect) regression line
text(2.9, 2.7, "d", font=3, cex=3)
text(2.9, 1.4, "d'", font=3, cex=3)
#legend("topleft", lty=c(1,2), legend=c("c (total effect)", "c' (direct effect)"), bty="n")
dev.off()

detach(package:lavaan)
```
Bully-victimization increased the predictive power of delusional-like experiences, increasing its parameter estimates. Suppressor.
## Check suppresson
```{r cronbach dataframes, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

check <- lm(ple_score ~ PA,  data = df2)
summary(check)

check2 <- lm(ple_score ~ PA + bully_score, data =  df2)
summary(check2)

check3 <- lm(ple_score ~ PA + as_index, data =  df2)
summary(check3)

check4 <- lm(ple_score ~ PA + as_index + bully_score, data =  df2)
summary(check4)


cor <- data.frame(latent_estimates$DLE, latent_estimates$BV,latent_estimates$AT,df2$PA)
cor(cor, method = "pearson")

library(ppcor)
pcor(cor)

## Partial correlation is lower than zero-order correlation indicating no suppression effects. May be due to facillitation rather than suppression.(not just picking up errors)

```






## Cronbach's Alpha Calculations

### Create Unique Dataframes

```{r cronbach dataframes, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Parse one score per ppt
ca <- df %>% 
  group_by(id) %>%
  distinct(id, .keep_all = TRUE) # unique score per ppts
```

#### Austic Traits Dataframe
```{r as dataframes, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Social skills scale 

trait_1 <- ca$TRAIT_1 #create variables to remove id
trait_4 <- ca$TRAIT_4
trait_7 <- ca$TRAIT_7
trait_9 <- ca$TRAIT_9
trait_11<- ca$TRAIT_11

ca_social_skills <- data.frame(trait_1, trait_4, trait_7,
                               trait_9, trait_11) # create new dataframe

## Attention switching scale

trait_2 <- ca$TRAIT_2
trait_5 <- ca$TRAIT_5
trait_10 <- ca$TRAIT_10
trait_12 <- ca$TRAIT_12


ca_attention_switching <- data.frame(trait_2, trait_5, trait_10, trait_12)

## Communication scale
trait_3 <- ca$TRAIT_3
trait_6 <- ca$TRAIT_6
trait_8 <- ca$TRAIT_8

ca_communication <- data.frame(trait_3, trait_6, trait_8)

## Total survey
ca_as_total <- data.frame(trait_3, trait_6, trait_8,trait_2,
                          trait_5,trait_10, trait_12, trait_1,
                          trait_4, trait_7, trait_9, trait_11)
```

#### Psychotic_like Experiences Dataframes

```{r ple dataframes, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Persecutary ideation scale

ple_1 <- ca$PLE_1
ple_3 <- ca$PLE_3
ple_4 <- ca$PLE_4
ple_5 <- ca$PLE_5
ple_6 <- ca$PLE_6

ca_pers_idea <- data.frame(ple_1, ple_3, ple_4, ple_5, ple_6)


## Bizzare experiences scale

ple_2 <- ca$PLE_2
ple_7 <- ca$PLE_7
ple_9 <- ca$PLE_9
ple_11 <- ca$PLE_11
ple_12 <- ca$PLE_12
ple_13 <- ca$PLE_13
ple_15 <- ca$PLE_15

ca_bizz_exp <- data.frame(ple_2, ple_7,ple_9,ple_11,
                          ple_12, ple_13, ple_15)

## Perceptual abnormalities scale

ple_8 <- ca$PLE_8
ple_10 <- ca$PLE_10
ple_14 <- ca$PLE_14

ca_perc_abn <- data.frame(ple_8, ple_10, ple_14)

## Total survey 

ca_ple_total <- data.frame(ple_2, ple_7,ple_9,ple_11, ple_12,
                           ple_13, ple_15, ple_8, ple_10, ple_14,
                           ple_1, ple_3, ple_4, ple_5, ple_6)

## Total without perceptual abnormalities scale

ca_ple_be_pi <- data.frame(ple_2, ple_7,ple_9,ple_11, ple_12,
                           ple_13, ple_15, ple_1, ple_3, ple_4, 
                           ple_5, ple_6)

```

#### Bullying Dataframes

```{r bully dataframes, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Total traditional victimisation bullying survey

tbv_1 <- ca$TBV.1_1
tbv_2 <- ca$TBV.1_2
tbv_3 <- ca$TBV.1_3
tbv_4 <- ca$TBV.1_4
tbv_5 <- ca$TBV.1_5
tbv_6 <- ca$TBV.1_6
tbv_7 <- ca$TBV.1_7
tbv_8 <- ca$TBV.1_8

ca_tbv_total <- data.frame(tbv_1, tbv_2,tbv_3, tbv_4, tbv_5, 
                           tbv_6, tbv_7, tbv_8)

## Total Cyber bullying victimisation survey

cbv_1 <- ca$CBV.1_1
cbv_2 <- ca$CBV.1_2
cbv_3 <- ca$CBV.1_3
cbv_4 <- ca$CBV.1_4
cbv_5 <- ca$CBV.1_5
cbv_6 <- ca$CBV.1_6
cbv_7 <- ca$CBV.1_7
cbv_8 <- ca$CBV.1_8
cbv_9 <- ca$CBV.1_9
cbv_10 <- ca$CBV.1_10
cbv_11 <- ca$CBV.1_11

ca_cbv_total <- data.frame(cbv_1, cbv_2,cbv_3, cbv_4, cbv_5, 
                           cbv_6, cbv_7, cbv_8, cbv_9, cbv_10,
                           cbv_11)

## Overall bullying

ca_bully_total <- data.frame(cbv_1, cbv_2,cbv_3, cbv_4, cbv_5, 
                             cbv_6, cbv_7, cbv_8, cbv_9, cbv_10,
                             cbv_11, tbv_1, tbv_2,tbv_3, tbv_4, 
                             tbv_5, tbv_6, tbv_7, tbv_8)
```

#### Performing Cronbach's Alpha
```{r ca test, warning = FALSE, echo = FALSE, message=FALSE, results= 'hide'}

## Temparily load package

library(ltm)


## Test autistic survey measures

cronbach.alpha(ca_social_skills, CI = TRUE,
               na.rm = TRUE) # Alpha .813 CI[0.768, 0.849]

cronbach.alpha(ca_attention_switching, CI = TRUE, 
               na.rm = TRUE) # Alpha 0.51 [0.397, 0.603]     

cronbach.alpha(ca_communication, CI = TRUE,
               na.rm = TRUE) # Alpha 0.791 [0.740, 0.837]

cronbach.alpha(ca_as_total, CI = TRUE,
               na.rm = TRUE) # Alpha 0.868 [0.843, 0.887]

## Test psychotic-like experience survey measures

cronbach.alpha(ca_pers_idea, CI = TRUE, 
               na.rm = TRUE) # Alpha 0.724 [0.648, 0.777] 

cronbach.alpha(ca_bizz_exp, CI = TRUE,
               na.rm = TRUE) # Alpha 0.693 [0.615, 0.751] 

cronbach.alpha(ca_perc_abn, CI = TRUE, 
               na.rm = TRUE) # Alpha 0.64 [0.512, 0.738]

cronbach.alpha(ca_ple_total, CI = TRUE, 
               na.rm = TRUE) # Alpha 0.822 [0.779, 0.854]

cronbach.alpha(ca_ple_be_pi, CI = TRUE,
               na.rm = TRUE) #Alpha 0.799 [0.757, 0.831]

## Test traditional and cyberbullyubg victimisation measures

cronbach.alpha(ca_tbv_total, CI = TRUE,
               na.rm = TRUE) #Alpha 0.872 [0.843, 0.895]

cronbach.alpha(ca_cbv_total, CI = TRUE,
               na.rm = TRUE) #Alpha 0.861 [0.799, 0.895]

cronbach.alpha(ca_bully_total, CI = TRUE,
               na.rm = TRUE) #Alpha 0.892 [0.865, 0.911]

## Remove package
detach(package:ltm)

```
